\title{\href{https://www.ual.es/estudios/grados/presentacion/plandeestudios/asignatura/4015/40154321?idioma=zh_CN}{Tecnologías Multimedia} - Study Guide - Milestone 10: Bit-rate Control through Quantization}

\maketitle

\section{Description}

At the hardware level, the samples of audio are usually represented
using \href{https://en.wikipedia.org/wiki/Pulse-code_modulation}{PCM
  (Pulse Code Modulation)}. In a PCM sample, the number of levels that
the signal can take depends on the
\href{https://en.wikipedia.org/wiki/Audio_bit_depth}{number of
  bits/sample} (16 bits in our case). This discussion is reffered only
to the
\href{https://en.wikipedia.org/wiki/Quantization_(signal_processing)}{quantization}
of digital signals.

Another key aspect to consider is that the processing that the
\href{https://en.wikipedia.org/wiki/Auditory_system}{Human Auditory
  System (HAS)} performs to understand audio signals has several
\href{https://en.wikipedia.org/wiki/Psychoacoustics}{sources of
  perceptual redundancy}. One of these sources of redundancy is the
\href{https://en.wikipedia.org/wiki/Equal-loudness_contour}{finite
  number of different volumen levels that we can recognize}~\cite{bosi2003intro}.

\href{https://en.wikipedia.org/wiki/Quantization_(signal_processing}{Scalar
  Quantization} (SQ) is the process of decreasing the number of levels
that a signal can
take~\cite{sayood2017introduction}. \href{https://en.wikipedia.org/wiki/Vector_quantization}{Vector
  Quantization} (VQ) is similar, but is applied tuples of
samples~\cite{foundationsofsignalprocessing}. SQ is used when the
samples are decorrelated because the
\href{https://en.wikipedia.org/wiki/Quantization_(signal_processing)#Rate%E2%80%93distortion_optimization}{coding
  efficiency} provided by VQ is
marginal~\cite{vetterli2014foundations}. In most lossy compression
systems, quantization is the only source of
distortion~\cite{taubman2012jpeg2000}.

Quantizers can be classified into
\href{https://en.wikipedia.org/wiki/Quantization_(signal_processing)#Mid-riser_and_mid-tread_uniform_quantizers}{uniform}
and
\href{https://nptel.ac.in/content/storage2/courses/117104069/chapter_5/5_5.html}{non-uniform}~\cite{sayood2017introduction,vetterli2014foundations}. An
uniform quantizator distributes the available representation levels
uniformely over the range of input values. Non-uniform quantizers use
different density of representation levels to those intervals of input
values that, for example, occur more often. Non-uniform quantizers can
also be classified into static and
\href{https://en.wikipedia.org/wiki/Adaptive_differential_pulse-code_modulation}{adaptive
  quantizers}. In the first case, the
\href{https://en.wikipedia.org/wiki/Probability_distribution}{distribution}
of the representation levels remain constant during the quantization
stage, and in the second case, the densities are adapted dinamically
to the characteristics of the input signal. In this milestone we will
use an
\href{https://en.wikipedia.org/wiki/Quantization_(signal_processing)#Dead-zone_quantizers}{uniform
  dead-zone scalar static quantizer}, which can be implemented
efficiently for digital signals and provides a good rate/distortion
trade-off. See \href{}{this Jupyter notebook} for more information.

The number of representation levels provided by a quantizer depends
basically on the so called quantization step, typically denoted by
$\Delta$. The higher the $\Delta$, the smaller the number of
representation levels, and therefore, the higher the distortion
generated by the quantization error. Obviously, the smaller the number
of representation levels, the smaller the output bit-rate. The control
of the bit-rate through the $\Delta$ values is a technique that
can be used in transmission systems to minimize the jitter and the
loss of packets.

\section{What you have to do?}

\begin{enumerate}
\item Create a class named \texttt{Quantization} that inherits from
  \texttt{Buffer} (the class implemented in the previous
  milestone). Store it in a file named \texttt{quantization.py}.
\item Override the methods $\mathtt{pack()}$ and $\mathtt{unpack()}$,
  which should implement:

\begin{pseudocode}{Quantization\_InterCom}{~}
  \PROCEDURE{Pack}{\mathtt{chunk\_number}, \mathtt{chunk}}
  \BEGIN
    \mathtt{quantized\_chunk} \GETS quantize(\mathtt{chunk})\\
    \mathtt{packed\_chunk} \GETS Buffer.pack(\mathtt{chunk_number}, \mathtt{quantized_chunk})\\
    return \mathtt{packed\_chunk}\\
  \END
  \ENDPROCEDURE
  \PROCEDURE{Unpack}{(\mathtt{packed\_chunk})}
  \BEGIN
    \mathtt{chunk\_number}, \mathtt{quantized\_chunk} \GETS Buffer.unpack(\mathtt{packed\_chunk})\\
    \mathtt{chunk} \GETS dequantize(\mathtt{quantized\_chunk})\\
    return \mathtt{chunk\_number}, \mathtt{chunk}
  \END
  \ENDPROCEDURE
\end{pseudocode}

\item Determine the quantization step $\Delta$ used in the
  $\mathtt{quantize()}$ method as a function of the throughput $B$ of
  the communication link. Take into consideration that in real
  intercommunication scenario, $B$ is time-variying and therefore,
  $\Delta = \mathtt{estimation}(B)$, where $\mathtt{estimation()}$
  should use the previously $B$ measured values. Such estimation
  should maximize the QoE.
\end{enumerate}

\section{Timming}

Please, finish this milestone at most in two weeks.

\section{Deliverables}

\begin{enumerate}
  \item The Python module named \texttt{quantization.py} stored in the
    \href{https://github.com/Tecnologias-multimedia/intercom}{root
      directory} of your \texttt{intercom}'s repo.
  \item A video demonstration showing that \texttt{quantization.py}
    adapts to the variations of the throughput of the link.
\end{enumerate}

\section{Resources}

\bibliography{JPEG2000,audio-coding,data-compression,signal-processing}
