% Emacs, this is -*-latex-*-

\input{../../definitions}
\title{Entropy Coding in the Subband Domain}

% ahora mismo, un chunk es la concatenación de 2048 (chunk_size)
% enteros de 24 bits. Los primeros 1024 bytes del plano de bytes más
% significativo son, generalmente aunque dependiendo del número de
% niveles de la DWT y del volumen del audio, los bits de signo de los
% coeficientes de la subbanda "mid" y los siguientes 1024 bytes son el
% bit de signo de los coeficientes de la subbanda side (que suelen ser
% más pequeños). En principio no debería existir correlación entre los
% coeficientes de la subbanda mid y los coeficientes de la subbanda
% side. Podría crearse un modelo probabilístico o un diccionaro
% independiente para cada subbanda, mid o side. Tampoco es esperable
% correlación temporal entre los enteros (signos) en este plano de
% bytes (tampoco entre los coeficientes). Por tanto, una codificación
% binaria de un sólo bit generaría 2048/8 = 256 bytes. DEFLATE
% actualmente está generando unos 180 bytes para q=256 y unos 350
% bytes para q=1. Con q=256 podría aplicarse un RLE en la parte signo
% del stereo, que se pierde (todos los coeficientes son 0).

% El plano de bytes intermedio

% Usando bytes, y tras quantizar, usar ZTC para acortar la longitud de la sequencia de bytes que finalmente codificamos con DEFLATE. Recordemos que si un coeficiente en la subbanda h^i es cero, entonces es basatante probable que todos sus descendientes sean cero también. Hay que tener en cuenta que tendremos que usar filtros DWT que no generen desplazamiento en la fase (filtros simétricos).

% Algoritmo: Input: enteros de 24 bits.

% 1. Sumar 32768. Todos los coeficientes ahora son positivos (mayores o iguales que 0). El 0 sólo se va a generar si 

% 2. 

\maketitle

\section{Description}

The entropic compressor used so far only takes into account that the
samples (or coefficients, depending on the number of layers used)
that, when compressed by byte planes, may contain a significant number
of zeros, thus increasing the compression rate. Furthermore, we know
that the coefficients tend towards zero as their frequency increases,
even without perceptual quantization.

\section{What you have to do?}

Implement an entropic coding algorithm that exploits the fact that,
starting from a certain coefficient (always beginning with the lowest
frequencies), and processing the coefficients by byte planes, from a
certain point onward, all the remaining bytes in that plane are
zero. To do this:

\begin{enumerate}
\item Make all coefficients positive by adding a (positive) offset
  (which must then be subtracted during decoding). The
  \texttt{00}-bytes of the binary representation of the positive
  coefficients should continue being \texttt{00}-bytes after this
  addition.
\item Using the byte \texttt{FF} as a zero-escape marker, implement this
  ``byte stuffing'' tecnique:
\begin{verbatim}
Example	             Original Data	Stuffed/Encoded Data
-------------------- ------------------ --------------------
Normal data	     [05, 12, 08]       [05, 12, 08]
Data has the marker  [05, FF, 08]       [05, FF, 01, 08]
End with zeros       [05, 00, 00, 00]   [05, FF, 00]
Stuffed data         [05, FF, 01, 08]   [05, FF, 01, 01, 08]
\end{verbatim}
Here, the stuffing byte is \texttt{01}. Obviously, the decoded data
exactly matches the original data.
\item Encode each byte-plane using DEFLATE.
\end{enumerate}

\section{Deliverables}

A new InterCom layer named \verb|zero_coding.py|. Write the code in a
notebook, along with an evaluation of this encoding technique.

\begin{comment}
  ESCAPE_BYTE = b'\xFF'
LITERAL_MARKER = b'\xFF\x01'
ZERO_PAD_MARKER = b'\xFF\x00'

def encode_zero_escape(data: bytes) -> bytes:
    """
    Encodes a byte array, escaping 0xFF and marking trailing zeros.
    """
    # 1. Strip all trailing zeros from the end of the data
    stripped_data = data.rstrip(b'\x00')
    
    # 2. Escape any naturally occurring 0xFF bytes in the remaining payload
    # This turns every 0xFF into 0xFF01
    encoded = stripped_data.replace(ESCAPE_BYTE, LITERAL_MARKER)
    
    # 3. If we stripped any zeros, append the "End with Zeros" marker
    if len(stripped_data) < len(data):
        encoded += ZERO_PAD_MARKER
        
    return encoded

def decode_zero_escape(encoded_data: bytes, target_length: int) -> bytes:
    """
    Decodes the byte array, restoring 0xFF and padding zeros if marked.
    Note: The decoder needs to know the original/target length to pad correctly.
    """
    # 1. Check if the payload ends with the "Pad with Zeros" marker
    if encoded_data.endswith(ZERO_PAD_MARKER):
        # Remove the marker from the end
        payload = encoded_data[:-2]
        
        # Unescape the literal 0xFFs, then pad the rest with zeros
        decoded = payload.replace(LITERAL_MARKER, ESCAPE_BYTE)
        return decoded.ljust(target_length, b'\x00')
        
    else:
        # No trailing zeros to pad, just unescape the literals
    return encoded_data.replace(LITERAL_MARKER, ESCAPE_BYTE)

# A 10-byte array with a natural 0xFF and trailing zeros
original_data = b'\x05\xFF\x10\x00\x00\x00\x00\x00\x00\x00'

# Encode it
encoded = encode_zero_escape(original_data)

print(f"Original Length: {len(original_data)}")
print(f"Original Data:   {original_data.hex(' ')}")
print(f"Encoded Length:  {len(encoded)}")
print(f"Encoded Data:    {encoded.hex(' ')}") 

# Decode it
decoded = decode_zero_escape(encoded, target_length=10)
print(f"\nDecoded Matches Original? {decoded == original_data}")
    
\end{comment}


\section{Resources}

\bibliography{maths,data_compression,DWT,audio_coding}

