% Emacs, this is -*-latex-*- 

\title{Tecnologías Multimedia \newline Study Guide}

\maketitle

\href{https://www.ual.es/estudios/grados/presentacion/plandeestudios/asignatura/4015/40154321}{TM
  (Tecnologías Multimedia)} is an optative subject of the Computer
Science Degree at the UAL (University of Almería).

\section{Methodology}
TM follows the PBL (Project-Based Learning) methodology. The students,
helped by the lecturer, develop a project during the classes. This
project is
\href{https://github.com/Tecnologias-multimedia/intercom}{InterCom}(municator),
a Python application that allows networked users to communicate, in
real-time, throught the Internet.

The project is developed as a sequence of milestones. The students, in
groups of up to 4 people, propose solutions for each milestone. Such
solutions are presented to the rest of the class, and each group gives
and receives a feedback from the rest of the class and also from the
lecturer. The best solutions are incorporated to the project, with the
objective of improve it.

\section{Milestones}

\begin{enumerate}
\item \href{https://tecnologias-multimedia.github.io/study_guide/framework/}{Framework}. % Día 0, hora 0.
\item \href{https://tecnologias-multimedia.github.io/study_guide/minimal/}{Meeting \texttt{minimal} InterCom}. % Hora 1. Sin presentación de alumnos.
\item \href{https://tecnologias-multimedia.github.io/study_guide/echo_cancellation/}{(TODO) Echo cancellation}.
\item \href{https://tecnologias-multimedia.github.io/study_guide/NAT_traversal/}{(TODO) NAT traversal}. % TODO
\item \href{https://tecnologias-multimedia.github.io/study_guide/latency/}{Hidding the network latency}. % 1 hora, sin presentación.
\item \href{https://tecnologias-multimedia.github.io/study_guide/BR_control/}{Bit-rate control}. % Estudiar otras propuestas. TODO
\item \href{https://tecnologias-multimedia.github.io/study_guide/threshold_of_hearing/}{(Partially TODO) Considering the threshold of hearing}.
%\item \href{https://tecnologias-multimedia.github.io/study_guide/simultaneous_masking/}{(TODO) Considering the simultaneous masking effect}.
\item \href{https://tecnologias-multimedia.github.io/study_guide/transform_coding/}{Transform coding for removing redundancy}.
%\item \href{https://tecnologias-multimedia.github.io/study_guide/perceptual_coding/}{Removing psychoacoustical redundancy}.
%\item \href{https://tecnologias-multimedia.github.io/study_guide/psychoacoustics/}{Removing psychoacoustical redundancy}.
\end{enumerate}

\section{Resources}

\bibliography{data-compression,signal-processing,DWT,linux,python,git,text-compression,maths,image-compression,JPEG2000,intercom,sound,networking,NAT,audio-coding}
