% Emacs, this is -*-latex-*-

\input{../definitions}
\title{\TM{} - Syllabus}
\title{\href{https://tecnologias-multimedia.github.io/}{Tecnologías Multimedia} \newline Syllabus}

\maketitle

\section{Contents and Scope}

Tecnologías Multimedia (TM) is an optative subject of the Computer
Science Degree at the UAL (University of Almería). TM is focused on
the development of real-time audio processing techniques,
specifically, encoding and transmission. the contents can be found
\href{https://tecnologias-multimedia.github.io/contents/}{here}.

\section{Main goals}

In this course we develop:
\begin{enumerate}
\item The ability to understand the environment of an organization and
  its needs in the field of information technologies and
  communications, in real time
  (\href{https://www.ual.es/application/files/8516/5061/5446/memoriavig-ing-informatica-4015.pdf}{TI1}).
\item The ability to conceive systems, applications and services based
  on network technologies, including Internet, web, e-commerce,
  multimedia, interactive services and mobile computing
  (\href{https://www.ual.es/application/files/8516/5061/5446/memoriavig-ing-informatica-4015.pdf}{TI6}).
\end{enumerate}

\section{Methodology}

TM follows the PBL (Project-Based Learning)
\href{http://portafirma.ual.es/pfirma/downloadReport/file?idDocument=4u61Ie5es2&idRequest=ZeBY35LlFa}{methodology}. The
students, helped by the lecturer, develop a project during the
classes. This project is
\href{https://github.com/Tecnologias-multimedia/intercom}{InterCom}(municator),
a Python application that allows networked users to communicate, in
real-time, throught the Internet.

The project is developed as a sequence of milestones. The students, in
groups of up to 4 people, propose solutions for each milestone. Such
solutions are presented to the rest of the class, and each group gives
and receives a feedback from the rest of the class and also from the
lecturer. The best solutions are incorporated to the project, with the
objective of improve it.

\section{Timetable and Attendance}

TM is organized in 14 lectures of 2 hours/session, and 14 practicals
of 2 hours/session, during 7 weeks (approx.). See the official
\href{https://www.ual.es/estudios/grados/presentacion/plandeestudios/asignatura/4015/40154321}{timetable}.

\section{Evaluation Criteria and Instruments}

TM is evaluated continuously, although there is the possibility of a
final evaluation (which would be carried out on the day scheduled for
the exam). The continuous evaluation of the students is carried out
from the work developed for each one of the milestones that are
presented in class.

The course evaluation process is the responsibility of the teacher and
follows the following rules:

\begin{enumerate}
\item Sufficiently in advance, the teacher will propose a series of
  milestones associated with the project that is developed in the
  subject. The teacher, both during the master classes and in the
  practical classes, will give all the necessary information so that
  the students can develop the milestones. Generally, all milestones
  can be developed in parallel, although there will be a proposed
  order of development.
\item The teacher will periodically receive information from the
  students about the degree of achievement of the milestones that are
  being developed, with the purpose of directing the development of
  the milestones. When the teacher check that the milestone has been
  correctly implemented, he will propose to all the groups of students
  to present their solutions publicly (in class).
\item The grade that students will receive, for groups of up to 4
  people, will depend on the order in which the solutions for the
  milestones have been reached. Specifically, the first group to
  achieve the milestone will receive the highest rating (10) and the
  rest of the groups, in temporal order of achievement of the
  milestone, will be rated with lower and lower grades, which will
  always be higher than 5 if the milestone has been properly
  developed.
\item All the students of the same group will receive the same grade,
  but internally (in the group) the students can modify the grade as
  long as the grade of the group remains constant.
\item The groups that do not present any solution will receive a score
  of 0 points. The groups that have not approved following these rules will
  be qualified as "Not Presented".
\end{enumerate}

This information is extended in the Teaching Guide
(\href{https://portafirma.ual.es/pfirma/downloadReport/file?idDocument=4Jp82utmug&idRequest=QY36GYcOZQ}{Español}/\href{https://portafirma.ual.es/pfirma/downloadReport/file?idDocument=Zcmom6qigD&idRequest=xXgueuk9oD}{English}).
